---
title: "About the package"

author: ""

date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{About the package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 10
)
```

## About the package 

CiteSource was developed in order to provide researchers the ability to examine the utility and efficacy of literature resources and search methodologies. The idea behind CiteSource is simply allowing users to deduplicate citation records, while maintaining customizable metadata about the citation.

Customizable metadata can include anything from a resource name (ex.Web of Science, LENS.org, PubMed), a method (database search, handsearching, citation snowballing), a variation used within a method (WoS string #1, Wos string #2, WoS string #3), a research phase (search, Ti/Ab screening, Full-text Screening), or a unique group of citations (benchmarking articles, articles from a previous review, articles with specific a affiliation). 

The CiteSource deduplication process is better described as record merging process due to the fact that the customizable metadata from duplicate records is maintained through the creation of a single, primary record. Beyond the merging of customizable metadata, the primary record is created by using the most complete metadata available in the DOI and Abstract fields. This means that when a primary record is created users know that it represents the most complete metatadata between the original records.

Once records are deduplicated, users are able to easily create plots and tables in order to answer specific questions or to simply explore the data to an effort to develop new hypotheses. Examples of questions may include how many unique records a specific source contributed or how the traditional method of developing a search string and applying it across databases fared against a new AI discovery tool in discovering relevant articles. Users may want to understand the overlap in records between two different search strings or evaluate the impact of including Google Scholar in a review. Before searching, a user may even develop a targeted search to better understand the topical coverage across databases they intend to search, and once the search has been developed, how particular source, string, or method performed in discovering benchmarking articles. 

## Customized metadata

CiteSource gives users the ability to add customized metadata to three different fields. These fields are "cite_source", "cite_string", and "cite_label", however, users are able to use these fields to enter any metadata associated with their inquiry. Currently only cite_string and cite_label are utilized for plot and table creation, further development that utilizes cite_string is planned.

## Installation
Use the following code to install CiteSource. Currently, CiteSource lives on GitHub and needs to be installed with the remotes package. In the future we will be working to make the package available via CRAN.

```{r eval = FALSE}
# Install the remotes packages to enable installation from GitHub
install.packages("remotes")
library(remotes)

# Install CiteSource
remotes::install_github("ESHackathon/CiteSource")
```

## Data reuse and exporting

## Development and license
Development of this project began as part of the Evidence Synthesis Hackathon and as part of Evidence Synthesis & Meta-Analysis in R Conference - ESMARConf 2022. to learn more about this awesome conference and hackathon please visit @ https://esmarconf.org/

CiteSource was created under the General Public License (>=v3) find more information on GPL @ https://www.gnu.org/licenses/gpl-3.0.html

## Shiny App
The CiteSource Shiny app is currently being developed and will be available @ https://estech.shinyapps.io/citesource/.
