---
title: "Plot Sandbox"

author: ""

date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot Sandbox}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 10,
  warning = FALSE
)
```
## About this vignette

This vignette is a bit of a sandbox for the experimentation with plots! Currently it is using the working example data (same data used in the "Source Analysis Across Screening Phases" vignette). This Vignette includes plots that are already built into CiteSource along with plots that we are exploring.

If you have any questions, feedback, ideas, etc. about this vignette or others be sure to check out our discussion board on github! https://github.com/ESHackathon/CiteSource/discussions/100

## 1. Installation of packages and loading libraries

```{r, results = FALSE, message=FALSE}
#Install the remotes packages to enable installation from GitHub
#install.packages("remotes")
#library(remotes)

#Install CiteSource
#remotes::install_github("ESHackathon/CiteSource")

#Load the necessary libraries
library(CiteSource)
library(dplyr)
library(ggplot2)
```

## 2. Import Reference Files and Custom Metadata

```{r}
#Import citation files from a folder
citation_files <- list.files(path = file.path("../vignettes/working_example_data"), pattern = "\\.ris", full.names = TRUE)

#Print citation_files to double check the order in which R imported our files.
citation_files
```


## 3. Tagging Sources & Deduplication

For this vignette we'll be using the same sources from the "Source Analysis Across Screening Phases" vignette as it's one of the larger sets of citations.

```{r}
# Set the path to the directory containing the citation files
file_path <- "../vignettes/working_example_data/"

metadata_tbl <- tibble::tribble(
  ~files,           ~cite_sources, ~cite_labels, 
   "AGRIS.ris",      "AGRIS",       "search",    
   "CAB.ris",        "CAB",         "search",    
   "EconLit.ris",    "EconLit",     "search",    
   "Final.ris",       NA,           "Final",     
   "GreenFile.ris",  "GreenFile",   "search",    
   "McK.ris",        "Method1",     "search",    
   "RM.ris",         "Method2",     "search",    
   "TiAb.ris",        NA,           "Screened",  
   "WoS_early.ris",  "WoS",         "search",    
   "WoS_later.ris",  "WoS",         "search"
) %>% 

dplyr::mutate(files = paste0(file_path, files))
	
citations <- read_citations(metadata = metadata_tbl)

# Deduplication & Identifying Crossover Records
unique_citations <- dedup_citations(citations)
#> formatting data...
#> identifying potential duplicates...
#> identified duplicates!
#> merging citations...

# Count number of unique and non-unique citations from different sources and labels
n_unique <- count_unique(unique_citations)

# Create dataframe indicating occurrence of records across sources
source_comparison <- compare_sources(unique_citations, comp_type = "sources")

```

### NEW TABLE Experimentation
```{r}
#Create a table to show the difference between the initial number of records uploaded for each source, the number Database-Specific Unique Records (D-SUR) which is the count after internal deduplication, and the number of unique records that source provided after deduplication across all sources.

library(DT)

# Function to count occurrences of each database
count_databases <- function(df, db_colname) {
  db_counts <- df %>%
    dplyr::pull(!!sym(db_colname)) %>%
    strsplit(", ") %>%
    lapply(unique) %>%  # Remove duplicates within each cell
    unlist() %>%
    trimws() %>%
    table() %>%
    as.data.frame() %>%
    dplyr::rename_with(~ gsub(" ", "_", .x, fixed = TRUE))
  
  colnames(db_counts) <- c("Source", "count")
  
  return(db_counts)
}

# Calculate D-SUR count for each database
d_sur_count <- count_databases(unique_citations, "cite_source")

# Count the citations in the initial citations data and remove blank rows
initial_citations_count <- count_databases(citations, "cite_source")

# Count the citations in the n_unique data, remove blank rows, and exclude rows without 'search' in cite_label
n_unique_citations_count <- n_unique %>%
  dplyr::filter(cite_label == "search") %>%
  dplyr::group_by(cite_source) %>%
  dplyr::summarise(Unique_records = sum(unique)) %>%
  dplyr::filter(cite_source != "") %>%
  dplyr::arrange(cite_source) %>%
  dplyr::rename(Source = cite_source)

# Combine the counts into one table
citation_counts <- dplyr::left_join(initial_citations_count, d_sur_count, by = "Source", suffix = c("_initial", "_D_SUR")) %>%
  dplyr::left_join(n_unique_citations_count, by = "Source", suffix = c("", "_Unique"))

# Update the column names
colnames(citation_counts) <- c("Source", "Records Imported", "D-SUR", "Unique records")

# Display the table using DT
datatable(citation_counts)

```
### Pull unique records from each source
```{r}
#Get unique records from each source and add bibliographic data
unique_AGRIS <- n_unique %>% 
  dplyr::filter(cite_source=="AGRIS", unique == TRUE) %>%
  inner_join(unique_citations, by = "duplicate_id")

unique_CAB <- n_unique %>% 
  dplyr::filter(cite_source=="CAB", unique == TRUE) %>%
  inner_join(unique_citations, by = "duplicate_id")

unique_EconLit <- n_unique %>% 
  dplyr::filter(cite_source=="EconLit", unique == TRUE) %>%
  inner_join(unique_citations, by = "duplicate_id")

unique_WoS <- n_unique %>% 
  dplyr::filter(cite_source=="WoS", unique == TRUE) %>%
  inner_join(unique_citations, by = "duplicate_id")

unique_Method1 <- n_unique %>% 
  dplyr::filter(cite_source=="Method1", unique == TRUE) %>%
  inner_join(unique_citations, by = "duplicate_id")

unique_Method2 <- n_unique %>% 
  dplyr::filter(cite_source=="Method2", unique == TRUE) %>%
  inner_join(unique_citations, by = "duplicate_id")

all_unique <- bind_rows(unique_AGRIS, unique_CAB, unique_EconLit,
                        unique_WoS, unique_Method1, unique_Method2)

```


## 4.PLOTS
### Heatmaps
```{r}
my_heatmap <- plot_source_overlap_heatmap(source_comparison)

my_heatmap

# Plot overlap as a heatmap matrix as percentage

my_heatmap_percent <- plot_source_overlap_heatmap(source_comparison, plot_type = "percentages")

my_heatmap_percent

```

### Upset plots
```{r}
# Citation duplication across multitiple resources

my_upset_plot <- plot_source_overlap_upset(source_comparison, decreasing = c(TRUE, TRUE))

my_upset_plot
```


### Unique/Crossover Bar Plots
```{r}
# Unique vs. Duplicate sources across Search/TiAb Screening/Final Inclusion

my_contributions <- plot_contributions(n_unique,
  center = TRUE,
  bar_order = c("search", "Screened", "Final")
)

my_contributions
```

### Publication year bar plots
```{r}

# Filter the "all_unique" dataset to only include records with a valid year
# Create a new column to represent the year
# Group the data by source and year and calculate the count of records for each group
year_data <- all_unique %>% 
  dplyr::mutate(year = as.numeric(year)) %>%  # Convert "year" column to numeric format
  dplyr::filter(!is.na(year) & year != "") %>%  # Remove rows where "year" is missing or empty
  group_by(cite_source.x, year) %>%  # Group data by "cite_source.x" and "year"
  summarise(count = n(), .groups = "drop")  # Calculate the number of records in each group

# Create a bar plot to visualize the filtered and grouped data
unique_yearplot <- ggplot(year_data, aes(x=year, y=count, fill = cite_source.x)) +
  geom_bar(position = "stack", stat = "identity") +  # Create a stacked bar plot with actual count values
  scale_x_continuous(breaks = seq(min(year_data$year), max(year_data$year), by = 5)) +  # Set x-axis breaks to every 5 years
  coord_cartesian(ylim = c(0, 950)) +  # Set y-axis limits to ensure all bars are visible
  xlab("Publication year") +  # Add x-axis label
  ylab("Unique records") +  # Add y-axis label
  labs(fill = "Source")  # Add legend label for the fill color

unique_yearplot
```

### interactive publication year bar plot
```{r}
# Creating an interactive plot with plotly
library(plotly)

unique_year_funplot <- ggplotly(unique_yearplot, tooltip = c("cite_source.x", "year", "count"), dynamicTicks = TRUE)

unique_year_funplot
```

### Sankey plot idea
```{r}
#further work needed to integrate Unique_citations into this plot, as this is very manual, but great as an example as to how the three custom fields can be visualized. In this case Strings/Sources/Timeline (Inclusion/Exclusion)

library(networkD3)

# Define nodes
nodes <- data.frame(
  name = c("AG 1", "AG 2", "AGRIS", 
           "CAB 1", "CAB 2", "CAB",
           "EL 1", "EL 2", "EconLit", 
           "WoS 1", "WoS 2", "WoS", 
           "Method 1", "Method 2", 
           "TI/AB Included", "Final Included", "Excluded", "NA"),
  id = 0:17
)

# Define links
links <- data.frame(
  source = c(0, 1, 3, 4, 6, 7, 9, 10),
  target = c(2, 2, 5, 5, 8, 8, 11, 11),
  value = c(100, 150, 100, 150, 100, 150, 100, 100)
)

# Add links for the second column nodes (including Method 1 and Method 2) flowing into TI/AB Included and Excluded
links <- rbind(links, data.frame(
  source = c(2, 5, 8, 11, 2, 5, 8, 11, 12, 12, 13, 13),
  target = c(14, 14, 14, 14, 16, 16, 16, 16, 14, 16, 14, 16),
  value = c(50, 50, 50, 50, 200, 200, 200, 150, 50, 50, 50, 50)
))

# Add links for TI/AB Included flowing into Final Included and Excluded
links <- rbind(links, data.frame(
  source = c(14),
  target = c(15),
  value = c(100)
))

links <- rbind(links, data.frame(
  source = c(14),
  target = c(16),
  value = c(200)
))

nodes <- rbind(nodes, data.frame(
  name = c("", ""),
  id = 17:18
))

links <- rbind(links, data.frame(
  source = c(17, 17),
  target = c(12, 13),
  value = c(NA, NA)
))



# Create Sankey diagram
sankey<-sankeyNetwork(
  Links = links,
  Nodes = nodes,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "name",
  units = "",
  fontSize = 12,
  nodeWidth = 20,
  iterations = 0
)
sankey
```

## 5. TABLES
### Unique Journal Count per Source 
```{r}
library(gt)
#Exploritory Plot - The following table and functions are not yet integrated into CiteSource

# Create a function to group by journal, count unique citations, and arrange in descending order
# This function takes a data frame as input, and returns a processed data frame
process_journals <- function(data) {
  data %>%
    group_by(journal) %>%          # Group the data by journal
    summarise(count = n()) %>%
    dplyr::filter(journal != "") %>%  # Filter out any blank journal titles
    arrange(desc(count))             # Arrange the journals in descending order of count
}

# Process Journal Titles
AGRIS_journals <- process_journals(unique_AGRIS)
CAB_journals <- process_journals(unique_CAB) 
EconLit_journals <- process_journals(unique_EconLit) 
WoS_journals <- process_journals(unique_WoS)
Method1_journals <- process_journals(unique_Method1)
Method2_journals <- process_journals(unique_Method2)

# Create a function to display a gt table
# This function takes a data frame and a caption as input, and returns a gt table
display_gt_table <- function(data, caption) {
  data %>%
    gt() %>%
    tab_header(title = caption) %>%  # Set the table title to the provided caption
    cols_label(journal = "Journal", count = "Unique Citations") %>%  # Set column labels
    tab_options(
      table.width = px(600),
      table.font.size = px(12),
      heading.title.font.size = px(16),
      heading.subtitle.font.size = px(14)
    )                                   # Set table styling options
}
```

### Top 10 journals for each source
```{r}
display_gt_table(CAB_journals[1:10, ], "Top 10 Journals by Unique Citations CAB")
```
```{r}
display_gt_table(EconLit_journals[1:10, ], "Top 10 Journals by Unique Citations EconLit")
```
```{r}
display_gt_table(WoS_journals[1:10, ], "Top 10 Journals by Unique Citations WoS")
```
```{r}
display_gt_table(Method1_journals[1:10, ], "Top 10 Journals by Unique Citations Method 1")
```
```{r}
display_gt_table(Method2_journals[1:10, ], "Top 10 Journals by Unique Citations Method 2")
```

### Citation Summary Table
```{r}
# Citation summary table

citation_summary_table(unique_citations, screening_label = c("Screened", "Final"))

```

### Citation Record Table
```{r}
# Citation Record Table

unique_citations %>%
  dplyr::filter(stringr::str_detect(cite_label, "Final")) %>%
  record_level_table(return = "DT")

```


